{\fontfamily{ptm}\selectfont Linear Convergence of Stochastic Nesterov's Accelerated Proximal Gradient method under Interpolation Hypothesis}

### 2025-07-09 Notebook LaTEX/SAGD Under Interpolation/main.tex
>w<, changes

\section{Nesterov's Accelerated Gradient}
    \subsection{In preparations}

@@

        \subsubsection{Basic definitions}

@@

        \subsubsection{Properties of function}
            \begin{definition}[semi quasi strongly convex function \textcolor{red}{NEW}]\label{def:sq-scnvx}
                A function $F: \RR^n \rightarrow \overline \RR$ is a semi quasi strongly convex function, abbreviated as ``SQ-SCNVX'' with respective to a linear mapping $A \in \RR^{m \times n}$ if $F - \frac{1}{2}\Vert Ax\Vert^2$ is a convex function. 

@@

                Any $\mu \ge 0$ strongly convex function is QS-SCNVX with $A = \sqrt{\mu}I$. 

@@

                One feature of a QS-SCNVX function is that it doesn't have a unique minimizer which differs it from strong convexity. 

@@

@@

            \begin{theorem}[quasi jesen inequality \textcolor{red}{NEW}]\label{thm:sq-scnvx-equiv}
                A function $F$ is semi quasi strongly convex with $A \in \RR^{m \times n}$ (Definition \ref{def:sq-scnvx}) if and only if, for all $x, y \in \RR^n$ and, $\lambda \in [0, 1]$ it satisfies the inequality: 
            \begin{definition}[smoothness and strong convexity with seminorm \textcolor{red}{NEW}]\label{def:seminorm-smooth-scnvx}
                Let $f:\RR^n \rightarrow \RR$ be a differentiable function on $\reli\dom f$. 
                Let $m \in \N$. 
                Let $A: \RR^{m\times n}$ and $b \in \RR^n$ be arbitrary. 
                Then, function $h: \RR^m \rightarrow \RR :=  x\mapsto f(Ax - b)$ is relatively smooth and, relatively strongly convex with respect to the function $x \mapsto (1/2)\Vert A x\Vert^2$: 
                    \frac{\mu}{2}\Vert Ax - Ay\Vert^2 \le 
                    D_f(x, y) \le \frac{L}{2}\Vert Ax - Ay\Vert^2. 
                The definition exchanged the $\Vert \cdot\Vert^2$ for a seminorm squared: $x \mapsto \Vert Ax\Vert^2$ with some $A: \RR^m \rightarrow \RR^n$. 
            The following theorem classifies a class of semi quasi strongly convex function. 

@@

@@

            We build the proximal gradient inequality for the class of SQ-SCNVX functions. 
        \subsubsection{Important inequalities}

@@

                Let $F = f + g$ and in addition assume that the set of minimizers $X^+ := \argmin_{x}F(x)$ is non-empty. 
            \begin{assumption}[seminorm smooth plus non-smooth]\label{ass:snorm-smth-p-nsmth}
                Let $F = f + g$. Assume that: 
                    \item $f: \RR^m \rightarrow \overline \RR$ is differentiable and, it satisfies Definition \ref{def:seminorm-smooth-scnvx} with $L > \mu \ge 0$ and $A \in \RR^{m \times n}$. 

@@

            \begin{theorem}[proximal gradient inequality with SQ-SCNVX \textcolor{red}{NEW}]\label{thm:pg-ineq-sq-scnvx}

@@

                Then, there exists some $B \ge 0$ such that $D_f(x^+, x) \le \frac{B}{2}\Vert x - x^+\Vert^2$ and, for all $z \in \RR^m, \eta \in \RR$ it satisfies the inequality: 

@@

                    + \frac{\eta}{2}\Vert z - x\Vert^2 

@@

                    + \frac{B - \eta}{2}\Vert z - x\Vert^2
                    - \frac{B}{2}\Vert z - x^+\Vert^2 . 

@@

                    &= F(z) - F(x^+)
                        \frac{\eta}{2}\Vert z - x\Vert^2 

@@

                    + \frac{B - \eta}{2}\Vert z - x\Vert^2
                    - \frac{B}{2}\Vert z - x^+\Vert^2 . 

@@

@@

@@

@@

@@

    \subsection{Stochastic accelerated proximal gradient}

@@

@@

                \item If, in addition, the sequence $\tilde \mu_k$ satisfies for all $k \ge 1$, $\frac{\tilde \mu_k}{L_{k - 1}} < L_{k - 1}/ L_k$, then the sequence also satisfies for all $k \ge 1$: $\alpha_k < 1$. 

@@

        \begin{lemma}[properties of the iterates]\label{lemma:snapg2-itrs-props}

@@

@@

    \subsection{Convergence rate of the algorithm under various circumstances}

@@

    \subsection{\textcolor{purple}{So, what to do next?}}
        Hi Arron would you like to add me for the co-authorship to continue this line of work and see how Nesterov's Accelerated Technique may work out for the stochastic gradient method? 
        These results are solid results but, they are still partial results and, below are the potential I foresee for this these ideas. 
        \begin{enumerate}
            \item Narrow down the sequence $\alpha_k$ and make sure that it can allow the quantity: 
            \begin{align*}
                \mathbb E_k\left[
                        \frac{(\alpha_k - 1)\mu^{(I_k)}\left(L_k\alpha_k - \mu^{(I_k)}\right)}{2\left(L_k - \mu^{(I_k)}\right)}
                    \right]\Vert x_{k - 1} - v_{k - 1} \Vert^2
            \end{align*}
            is negative, or at least bounded. I am not sure how this will work out, but I have some solid ideas around it. 
            \item Roll up the inequality in Theorem \ref{thm:snapg2-one-step} recursively and, determine the convergence rate through $\alpha_k$ that makes the previous item true. 
            In addition, I have the hunches that the convergence rate involves the variance of $\mu^{(I_k)}$ and, it will slower than the non-stochastic case of the algorithm. 
        \end{enumerate}
        For the future we can: 
        \begin{enumerate}
            \item Extend the definition of strong convexity to relative strong convexity with respect to a quasi-norm. This would extend interpolation hypothesis in Assumption \ref{ass:interp-hypothesis} where, even if $\mu > 0$, it doesn't mean that $F$ has a unique solution through strong convexity. This is entirely possible and appeared in the literatures before so, I can give you the words of confidence. 
            \item Show the convergence of the method for objective function based on quasi-strong convexity. This is a much weaker assumption it works well in practice for the common known problems in convex programming. 
        \end{enumerate}

>>>>

\section{In preparations}

@@

    \subsection{Basic definitions}

@@

    \subsection{Properties of functions, characterizations}
        The definitions are ordered from the weakest to strongest. 
        \begin{definition}[semi strongly convex function \textcolor{red}{NEW}]\label{def:semi-scnvx}
            A function $F: \RR^n \rightarrow \overline \RR$ is a semi strongly convex function, abbreviated as ``Semi-SCNVX'' with respective to a linear mapping $A \in \RR^{m \times n}$ if $F - \frac{1}{2}\Vert Ax\Vert^2$ is a convex function. 

@@

            Any $\mu \ge 0$ strongly convex function is Semi-SCNVX with $A = \sqrt{\mu}I$. 

@@

            One feature of a Semi-SCNVX function is that it doesn't have a unique minimizer which differs it from strong convexity. 

@@

        \begin{definition}[semi relative smoothness and Semi-SCNVX \textcolor{red}{NEW}]\label{def:seminorm-smooth-scnvx}
            Let $m, n\in \N$ be a natural numbers. 
            Let $f:\RR^m \rightarrow \RR$ be a differentiable function and full domain. 
            If there exists $A_1: \RR^{m\times n}$ and, $A_2 \in \RR^{m \times n}$ matrices such that it satisfies:
                \frac{1}{2}\Vert A_1x - A_1y\Vert^2 \le 
                D_f(x, y) \le \frac{1}{2}\Vert A_2x - A_2y\Vert^2. 
            Then, we call this function is semi-relative smooth with respect to $A_1$, and semi convex to $A_2$
            The definition exchanged the $\Vert \cdot\Vert^2$ for a seminorm squared: $x \mapsto \Vert A_1x\Vert^2$ with some $A: \RR^m \rightarrow \RR^n$ for the definition of relative smoothness and relative strong convexity. 
            Obviously, it has $\Vert A_1x\Vert \le \Vert A_2x\Vert$ for all $x\in \RR^m$, which further implies that $\ker A_1\supseteq\ker A_2$. 
        \begin{theorem}[semi Jesen inequality \textcolor{red}{NEW}]\label{thm:semi-scnvx-equiv}
            A function $F$ is Semi-SNCVX with $A \in \RR^{m \times n}$ (Definition \ref{def:semi-scnvx}) if and only if, for all $x, y \in \RR^n$ and, $\lambda \in [0, 1]$ it satisfies the inequality: 

@@

        The following theorem classifies a class of semi strongly convex function. 

@@

            Therefore, it satisfies Definition \ref{def:semi-scnvx} with $A_1 = \sqrt{L}A, A_2 = \sqrt{\mu}A$. 

@@

        We build the proximal gradient inequality for the class of Semi-SCNVX functions. 
    \subsection{Important inequalities}
            The following assumptions are ordered from weakest to strongest. 
            All of them are some of smooth with a non-smooth function. 
            \begin{assumption}[affine S-CNVX composition plus non-smooth]\label{ass:snorm-smth-p-nsmth}
                Let $A \in \RR^{m\times n}, b \in \RR^n$
                Let $F(x): \RR^{m} \mapsto \overline\RR := x \mapsto f(x) + g(x)$. 
                Assume that: 
                    \item $f:\RR^m \rightarrow \RR := x \mapsto h(Ax - b)$ where $h:\RR^n\rightarrow \RR$ is $L$ Lipschitz smooth and, $\mu\ge 0$ strongly convex, then it satisfies Theorem \ref{thm:smooth-aff-sq-scnvs-fxn} with $L > \mu \ge 0$ and $A \in \RR^{m \times n}$. 

@@

                Let $F = f + g$ satisfies \ref{ass:smooth-plus-nonsmooth} and in addition assume that the set of minimizers $X^+ := \argmin_{x}F(x)$ is non-empty. 

@@

            The following theorem attempts to generalize Theorem \ref{thm:pg-ineq}. 
            \begin{theorem}[proximal gradient inequality with semi-scnvx \textcolor{red}{NEW}]\label{thm:pg-ineq-semi-scnvx}

@@

                Let $\sigma_{\min}(A)$ denotes the minimum non-zero singular value of $A$ in absolute value. 
                Let $\Pi = \Pi_{\ker A}$ be the linear operator that project onto the kernel of $A$. 
                Then, there exists some $B \ge 0$ such that $D_f(x^+, x) \le \frac{B}{2}\Vert x - x^+\Vert^2$ and, for all $z \in \RR^m$ it satisfies the inequality: 

@@

@@

                    &\le F(z) - F(x^+)
                    + \frac{B - \sigma_{\min}(A)^2\mu}{2}\Vert (I - \Pi)(z - x) \Vert^2
                        + \frac{B}{2}\Vert \Pi(z - x)\Vert^2. 

@@

                    F(z) - F(x^+) 

@@

                    - \frac{B}{2}\Vert z - x^+\Vert^2. 

@@

                Since $\Pi = \Pi_{\ker A}$ to be the projection onto the kernel of matrix $A$, $z - x$ can be written as $\Pi(z - x) + (z - x) - \Pi(z - x)$ and so: 
                    & \Pi(z - x) \perp (z - x) - \Pi(z - x)
                    \implies & 
                    \Vert z - x\Vert^2 = \Vert \Pi(z - x) \Vert^2 + \Vert (z - x) - \Pi(z - x)\Vert^2
                Let $\sigma_{\min}(A)$ denotes the minimum non-zero singular value of $A$ in absolute value, then it has: 
                    \Vert A(x - z)\Vert^2 &= \Vert A( \Pi(z - x)  + (z - x) - \Pi(z - x))\Vert^2
                    &= \Vert A((z - x) - \Pi(z - x))\Vert^2
                    &\ge \sigma_{\min}(A)^2\Vert (z - x) - \Pi(z - x)\Vert^2.
                Using two of the above results to keep simplifying the inequality, it yields: 
                    F(z) - F(x^+)
                    F(z) - F(x^+)
                    + \frac{B - \sigma_{\min}(A)^2\mu}{2}\Vert (z - x) - \Pi(z - x) \Vert^2
                        + \frac{B}{2}\Vert \Pi(z - x)\Vert^2. 
                    F(z) - F(x^+)
                    + \frac{B - \sigma_{\min}(A)^2\mu}{2}\Vert (I - \Pi)(z - x) \Vert^2
                        + \frac{B}{2}\Vert \Pi(z - x)\Vert^2. 
                Last equality we used the fact that $\Pi$ is an linear operator as well. 

@@

                When $\ker A = \{\mathbf 0\}$, this theorem is equivalent to Theorem \ref{thm:pg-ineq} but with $\mu$ being $\sigma_{\min}(A)$ instead. 
            The following theorem attempts to generalize Theorem \ref{thm:semi-scnvx-equiv} for relatively smooth plus non-smooth function. 

@@

                Let $\Pi = \Pi_{\ker A}$ be the projection onto the kernel of $A$. 
                Let $\sigma_{\min}(A)$ denote the smallest non-zero singular value of $A$ in absolute value. 
                Then, for all $x, y \in \RR^m$ and, $\lambda \in [0, 1]$ it satisfies the inequality: 
                    F(\lambda z + (1 - \lambda)y) &\le \lambda F(z) + (1 - \lambda)F(x) - \frac{\sigma_{\min}(A)^2\lambda(1 - \lambda)\mu}{2}\Vert (I - \Pi)(x - y)\Vert^2. 

@@

                $f$ satisfies Definition \ref{def:seminorm-smooth-scnvx} with $L > \mu, A$, so for all $x, y \in \RR^m$ it has 
                    0 &\le D_f(x, y) - \frac{\mu}{2}\Vert Ax - Ay\Vert^2 
                    \le \frac{L - \mu}{2}\Vert Ax - Ay\Vert^2. 
                Using some algebra (or equivalent some properties of Bregman divergence), it shows that the function $f - \mu/2 \Vert A(\cdot)\Vert^2$ is a convex function, therefore, $f + g - \mu/2\Vert A(\cdot)\Vert^2 = F - \frac{\mu}{2}\Vert A(\cdot)\Vert^2 = F - \frac{1}{2}\Vert \sqrt{\mu}A(\cdot)\Vert^2$ is also a convex function. 
                Applying Theorem \ref{thm:semi-scnvx-equiv} it has for all $z, x \in \RR^m$ and, $\lambda \in [0, 1]$ the inequality: 
                    F(\lambda z + (1 - \lambda)y) &\le \lambda F(z) + (1 - \lambda)F(x) - \frac{\lambda(1 - \lambda)\mu}{2}\Vert Ax - Ay\Vert^2. 
                Let $\Pi = \Pi_{\ker A}$ be the projection onto the kernel of $A$, then it has: 
                    \Vert Ax - Ay\Vert^2 &= \Vert A\Pi(x - y) + (I - \Pi)(x - y) \Vert^2
                    \Vert A\Pi(x - y) \Vert^2 + \Vert A(I - \Pi)(x - y) \Vert^2
                    &\ge 0 + \sigma_{\min}(A)^2\Vert (I - \Pi)(x - y) \Vert^2
                Therefore, it gives the weaker inequality: 
                    F(\lambda z + (1 - \lambda)y) &\le \lambda F(z) + (1 - \lambda)F(x) - \frac{\sigma_{\min}(A)^2\lambda(1 - \lambda)\mu}{2}\Vert (I - \Pi)(x - y)\Vert^2. 

@@

\section{Stochastic accelerated proximal gradient}
    In this section, we define the Stochastic accelerated Nesterov's Acceleration algorithm and, progressively build up the results to the one step convergence of the algorithm. 

@@

    \begin{assumption}[sum of many affine composite]\label{ass:sum-of-many-aff-comp}
        Let $A^{(1)}, A^{(2)}, \ldots A^{(n)}$ be a list of $\RR^{n\times m}$ matrices. 
        Suppose $F:\RR^n \rightarrow \overline \RR:= (1/n)\left(\sum_{i = 1}^{n} f_i(x) + g_i(x)\right)$. 
        Assume
            \item 

@@

            \item If, in addition, the sequence $\tilde \mu_k$ satisfies for all $k \ge 1$, $\frac{\tilde \mu_k}{L_{k - 1}} < L_{k - 1}/ L_k$, then the sequence strictly less than one and, for all $k \ge 1$: $\alpha_k \in (0, 1)$. 

@@

    \begin{lemma}[properties of the iterates \textcolor{red}{NEW}]\label{lemma:snapg2-itrs-props}

@@

    \end{proof}
    % --------------------------------------------------------------------------------------------------------------
    \begin{lemma}[SNAPG-V2 one step convergence prototype stage-I]\label{lemma:snagp2-one-step-s1-proto}

    \end{lemma}
    \begin{proof}

@@

\section{Convergence rate of the algorithm under various circumstances}

@@

\section{\textcolor{purple}{So, what to do next?}}
    Hi Arron would you like to add me for the co-authorship to continue this line of work and see how Nesterov's Accelerated Technique may work out for the stochastic gradient method? 
    These results are solid results but, they are still partial results and, below are the potential I foresee for this these ideas. 
    \begin{enumerate}
        \item Narrow down the sequence $\alpha_k$ and make sure that it can allow the quantity: 
        \begin{align*}
            \mathbb E_k\left[
                    \frac{(\alpha_k - 1)\mu^{(I_k)}\left(L_k\alpha_k - \mu^{(I_k)}\right)}{2\left(L_k - \mu^{(I_k)}\right)}
                \right]\Vert x_{k - 1} - v_{k - 1} \Vert^2
        \end{align*}
        is negative, or at least bounded. I am not sure how this will work out, but I have some solid ideas around it. 
        \item Roll up the inequality in Theorem \ref{thm:snapg2-one-step} recursively and, determine the convergence rate through $\alpha_k$ that makes the previous item true. 
        In addition, I have the hunches that the convergence rate involves the variance of $\mu^{(I_k)}$ and, it will slower than the non-stochastic case of the algorithm. 
    \end{enumerate}
    For the future we can: 
    \begin{enumerate}
        \item Extend the definition of strong convexity to relative strong convexity with respect to a quasi-norm. This would extend interpolation hypothesis in Assumption \ref{ass:interp-hypothesis} where, even if $\mu > 0$, it doesn't mean that $F$ has a unique solution through strong convexity. This is entirely possible and appeared in the literatures before so, I can give you the words of confidence. 
        \item Show the convergence of the method for objective function based on quasi-strong convexity. This is a much weaker assumption it works well in practice for the common known problems in convex programming. 
    \end{enumerate}
